Open-source landscape for 3D CSEM modelling

### 2020-04-27 manuscript/Manuscript.tex
Re-work up before Marlim; next Marlim

\imycom[DW]{
We will have to decide on a Journal. Current suggestions:\\
- GMD: Geoscientific Model Development\\
- SIG: Surveys in Geophysics\\
- SE: Solid Earth\\
- GJI: Geophysical Journal International\\
- GP: Geophysical Prospecting\\
- GEO: Geophysics\\
- CAG: Computers and Geosciences\\
I think \emph{Geoscientific Model Development} and \emph{Surveys in Geophysics}
are the favoured currently. I personally would be very interested in GMD. What
do you think? Also interesting would be the GJI, because the MT comparison was
published there, so it would be in the same place.
}

@@

The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of publications. The amount
of available numerical solutions can be overwhelming and is part of the reason
why there are hundreds of publications about the topic. There are the different
methods to solve Maxwell's equation, such as the IE method
\citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani,
CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the
differential equation (DE) method, for instance finite differences (FD)
\citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich,
CAG.13.Sommer}, finite elementes (FE) \citep{GJI.11.Schwarzbach, GEO.04.Commer,
GEO.12.daSilva, GJI.13.Puzyrev, GJI.13.Grayver, SEG.16.Zhang}, finite volume
(FV) \citep{EM.90.Madsen, ECP.07.Haber, GEO.14.Jahandari, PIER.01.Clemens,
GP.06.Mulder}. And these are just the most common ones.

@@

The biggest variety of all exists probably in the available solvers to solve
the system of linear equations; direct solvers \citep{GEO.09.Streich,
GEO.15.Grayver, GP.14.Chung, GEO.14.Jaysaval, SEG.15.Oh, GJI.18.Wang}, indirect
solvers \citep{GP.06.Mulder, GJI.15.Jaysaval} or a combination of both,
so-called hybrid solvers \citep{GEO.18.Liu}; the solvers often use
preconditioners such as the multigrid method \citep{SIAM.02.Aruliah,
GP.06.Mulder, GJI.16.Jaysaval}.

A very well written overview up to the year 2005 of the different approaches to
3D EM modelling is given by \cite{SG.05.Avdeev}. In the last 15 years the
publications with regards to 3D EM  modelling grew tremendously, driven
by the ever increasing computing power. Avdeev finished his review with the
following statement: «\emph{The most important challenge that faces the EM
community today is to convince software developers to put their 3-D EM forward
and inverse solutions into the public domain, at least after some time. This
would have a strong impact on the whole subject and the developers would
benefit from feedback regarding the real needs of the end-users.}»

It is this topic that we address, open-source 3D CSEM codes. Some major
exploration service companies as well as certain research consortia and
research groups have their 3D CSEM codes. But how about codes that are
open-source? Traditionally research code was often available upon request by
the author. Codes distributed that way often come with the request to not share
the code, or they often come without a license attached at all (which means
that they are not open-source). Also, at times there can be a significant
hurdle to install a code. However, the meaning of open-source evolved rapidly
in the last decades. Today, open-source generally not only means that a code is
available (with a proper license). It includes much more today, such as that
the code is hosted online, is under version control, has the possibility to
file issues and make pull requests to fix bugs. Well maintained codes often
have continuous integration that includes testing of the code base and online
hosted documentation. As such the term has shifted from purely open-source code
to development in the open and increasingly building a global community. There
are a number of projects within the realm of geophysical exploration, besides
the ones we present here, \todo{Others coming to your mind?}e.g. pyGIMLi
\citep{CAG.17.Rucker} and Fatiando \citep{JOSS.18.Uieda}. Newer \todo{wrong
term, think about it}programming environments  have also brought much
simplifications in the installation process and have most importantly
simplified it for code developers to make their codes easily available and
installable. All our presented codes are in the Python ecosystem, and each code
can be installed with a single command (which includes downloading and
installation).

A second topic we want to address is verification. New codes are generally
verified against analytical solutions. However, analytical and semi-analytical
solutions only exist for very simple models. There is no possibility to check
how good a computed response is, expect for comparing the result from different
modellers. If different modellers using different discretisation and
implementations of Maxwell's equations yield the same result it can give
confidence in the accuracy. Providing these two models and results from four
different codes should give a benchmark with which new codes can be compared
and validated.

It is worth mentioning that there are many more 3D EM codes developed in other
fields than geophysics such as communication systems (antennas, radar,
satellites), medical imaging. There are various open-source codes in these
fields \todo{Add at least three references} too, e.g., Elmer
(\href{http://www.csc.fi/elmer}{csc.fi/elmer}, couldn't find a publication,
just a poster). While these codes could potentially be used for the same goals
as presented here we restrict our review to codes purpose-built for geophysical
applications.

In the following section we introduce the codes under consideration. The
numerical result consider two cases. First a layered background with vertical
transverse isotropy containing three blocks, followed by a realistic, complex
model.

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}. All four
codes have their user-facing routines written in Python; all of them make heavy
use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy}
\citep{NM.20.Virtanen}. All are “modern” open-source projects, meaning that
they not only come with an open-source license, but they are also in an
online-hosted version-control system with feedback and tracking possibilities
(raising issues, filing pull requests), have extensive online documentation
including many examples, have continuous integration to some degree, and also
importantly are all easily installable, without any user-side compilations.

In the following a quick rundown of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the Maxwell's equation in its differential form (DE)
under the quasistatic or diffusive approximation, hence neglecting displacement
currents. The machines on which the different codes were run are listed in
Table~\ref{tbl:machines}, together with the responsible operator.

\tabl[btp]{codecomparison}{All codes solve Maxwell's equations in its
  differential form (DE) under the quasistatic approximation. Note that
  \emg3d is a solver on its own, while the other codes implement third-package
  solvers such as \texttt{PETSc}, \texttt{MUMPS}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
  \imycom[DW]{\texttt{PETSc} and \texttt{MUMPS} probably need citations too.}
  \imycom[DW]{PEC/PMC instead of Dirichlet/Neumann?}
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & Neumann; Dirichlet & PEC & PEC & Dirichlet; Neumann \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

\imycom[DW]{
Each code should be briefly summarized by three points:\\
- Brief intro\\
- Main selling point\\
- Planned features\\
I think in a first round \textbf{everyone should just write what they want to
write and think is important for their code}. In a second round we should then
homogenize it, such that all codes have more or less the same length and also
similar topic/layout.
}

@@

\texttt{FEniCS} \citep{CSE.15.Alnaes}

@@

The 3D CSEM modeller \emg3d is a multigrid \citep{CMMP.64.Fedorenko} solver for
3D electromagnetic diffusion following \cite{GP.06.Mulder}, with tri-axial
electrical anisotropy, isotropic electric permittivity, and isotropic magnetic
permeability. The matrix-free solver can be used as main solver or as
preconditioner for one of the Krylov subspace methods implemented in SciPy, and
the governing equations are discretized on a staggered grid by the
finite-integration technique \citep{AEU.77.Weiland}. The code is written
completely in Python using the NumPy and SciPy stacks, where the most time- and
memory-consuming parts are sped up through jitted (just-in-time compiled)
functions using Numba \citep{LLVM.15.Lam}. The code computes the electric field
due to an electric source in the frequency-domain. Implemented are functions to
obtain the magnetic field due to an electric source; to obtain the electric and
magnetic fields due to a magnetic source; and to obtain time-domain responses.

The multigrid method is characterized by almost linear scaling both in terms of
runtime (CPU) and memory (RAM) usage, and it is therefore a comparably
low-memory consumption solver. It is also minimal in terms of requirements,
only NumPy, SciPy, and Numba are required, with Python 3.7+. The current
development is focused on adding inversion capabilities. A further plan is also
to implement a hook in order to use \emg3d as a solver for \simpeg.

@@

@@

%
\tabl[btp]{machines}{List of computer and operating system on which the
  different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & PowerEdge R940 server; 144 Xeon Gold 6154 CPU @2.666 GHz; ~3 TB
            DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & Laptop; i7-6600U CPU@2.6 GHz x4; 16 GB of memory, Ubuntu 18.04
          & Dieter Werthmüller \\
  \petgem & Marenostrum4. Intel Xeon Platinum from Skylake generation; 2
            sockets Intel Xeon Platinum 8160 CPU with 24 cores each @2.10GHz
            for a total of 48 cores per node; 386 Gb DDR4 RAM per node; SuSE
            Linux Enterprise
          & Octavio Castillo-Reyes \\
  \simpeg & ???
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%

\clearpage  % TODO: Just for now while drafting, remove later!
\section{Verifications}

@@

\cite{GJI.13.Miensopust} published a review of two workshops dealing with the
verification of magnetotelluric forward and inversion codes, hence a similar
exercise to what we do here for CSEM codes. The block model is a derivation of
the \emph{Dublin Test Model 1}, a model from their first workshop. We took the
same layout of the blocks but adjusted the dimensions and resistivities to a
typical, marine CSEM problem, as shown in Figure~\ref{fig:block-model}. We
additionally added a layered background with vertical transverse isotropy
(VTI).
%
\plot*{block-model}{width=.6\textwidth}{
  Sketch of the block model. The layered model consists of an air layer, a
  water layer, a thin top-layer and then a thick background layer, and at the
  bottom a resistive basement layer. All three blocks are embedded in the thick
  background layer, which has VTI with $\lambda=\sqrt{2}$.
}
%

@@

Figure~\ref{fig:results-layered-0} shows in the top-row the inline responses
for offsets $r>0.5\,$km. As all codes compute the same we only show the
semi-analytical result for the real and imaginary amplitudes, and then in the
bottom row the relative error of the 3D codes. The results show that the codes
are literally yielding the same result, the relative error is for big parts
below 1\,\%, and for almost all parts below a few percent. Some differences are
due to the chosen discretizations. E.g., \emg3d is less precise close to the
200\,m long source. This could be improved by choosing smaller cell sizes, at
the cost of computation time. At large offsets, on the other hand, \emg3d seems
to do a very good job.
%
\plot*{results-layered-0}{width=.8\textwidth}{
   Comparison of the layered model for the inline receivers ($y=0\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   show the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

@@

%
\tabl[btp]{comp-layered}{Comparison of used CPU and RAM and of number of cells
  used to discretize the layered model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#Cells \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 & 24 & 0.0 & 0 \\
  \emg3d  & 0.0 &  1 & 0.0 & 0 \\
  \petgem & 0.0 & 48 & 0.0 & 0 \\
  \simpeg & 0.0 &    & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

Three resistive blocks are added in the 3.3\,km-thick, anisotropic background
layer, as shown in Figure~\ref{fig:block-model}. They have resistivities of
$\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines),
$\rho=100\,\ohmm$ (thin plate, South-East), $\rho=500\,\ohmm$ (cube,
North-West). As there are no analytical solutions for this model we show the
normalized difference between different codes, instead of the relative error,
where the normalized difference of two responses $p$ and $q$ is given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200 \frac{|p - q|}{|p| + |q|}\ .
  \label{eq:nrmsd}
\end{equation}
%
\imycom[DW]{NRMSD: Comparing all codes to all would make the
figures too cluttered, and I don't think it would add a lot insights. My idea
is therefore to compare \emg3d/\simpeg (the two FD ones), \custem/\petgem (the
two FE ones), and \emg3d/\custem (a cross-comparison). What do you think?}

@@

The difference becomes only significant at offsets larger than 8\,km. However,
this is due to a practical difference in the meshing: \textbf{Rewrite, Comments
Raphael} “\emph{Yes, cutting the layers has an influce at offsets greater
  distance X (see example \#2 thesis) - depending on the boundary mesh. I'm
  sure that's the reason. I could also add a version with a way larger
  horizontal layered earth extent, but this would require way more resources -
  unstructured meshes are just not competitive to regular grids for regular
geopmetries in such cases as discussed before}“

>>>>

\imycom[DW]{
We will have to decide on a Journal. Current suggestions:\\
- GMD: Geoscientific Model Development\\
- SIG: Surveys in Geophysics\\
- SE: Solid Earth\\
- GJI: Geophysical Journal International\\
- GP: Geophysical Prospecting\\
- GEO: Geophysics\\
- CAG: Computers and Geosciences\\
I think \emph{Geoscientific Model Development} and \emph{Surveys in Geophysics}
were the favourite ones when we discussed it ones. I personally would be very
interested in GMD. What do you think? But also interesting would be the GJI,
because the MT comparison was published there, so it would be in the same
place.
}

@@

The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of publications. The amount
of available numerical solutions can be overwhelming and is part of the reason
why there are hundreds of publications about the topic. There are the different
methods to solve Maxwell's equation, such as the IE method
\citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani,
CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the
differential equation (DE) method, for instance finite differences (FD)
\citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich,
CAG.13.Sommer}, and finite elementes (FE) \citep{GJI.11.Schwarzbach,
GEO.04.Commer, GEO.12.daSilva, GJI.13.Puzyrev, GJI.13.Grayver, SEG.16.Zhang},
finite volume (FV) \citep{EM.90.Madsen, ECP.07.Haber, GEO.14.Jahandari,
PIER.01.Clemens, GP.06.Mulder}. And these are just the most common ones.

@@

The biggest variety of all exists probably in the available solvers to solve
the system of linear equations; direct solvers \citep{GEO.09.Streich,
GEO.15.Grayver, GP.14.Chung, GEO.14.Jaysaval, SEG.15.Oh, GJI.18.Wang}, indirect
solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both,
so-called hybrid solvers \citep{GEO.18.Liu}; the solvers often use
preconditioners such as the multigrid method \citep{SIAM.02.Aruliah,
GP.06.Mulder, GJI.16.Jaysaval}.

A very well written overview up to the year 2005 of the different approaches to
3D EM modelling is given by \cite{SG.05.Avdeev}, and also by
\cite{SG.10.Borner}. In the last 15 years the publications with regards to 3D
EM  modelling grew tremendously, driven by the ever increasing computing power.
Avdeev finished his review with the following statement: «\emph{The most
important challenge that faces the EM community today is to convince software
developers to put their 3-D EM forward and inverse solutions into the public
domain, at least after some time. This would have a strong impact on the whole
subject and the developers would benefit from feedback regarding the real needs
of the end-users.}»

It is this topic that we address, open-source 3D CSEM codes. Some major
exploration service companies as well as certain research consortia and
research groups have their 3D CSEM codes. But how about codes that are
open-source? Research code was traditionally often available upon request by
the author. Codes distributed that way often come with the request to not share
the code, and they often come without a license attached at all (which means
that they are not open-source). Also, at times there can be significant hurdles
to install a code. However, the meaning of open-source evolved rapidly in the
last decades. Today, open-source generally not only means that a code is
available (with a proper license). The meaning includes much more these days,
such as that the code is hosted online, is under version control, has the
possibility to file issues and make pull requests to fix bugs. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation which often includes code documentation,
examples, and tutorials. As such the term has shifted from purely open-source
code to development in the open and increasingly building a global community.
There are a number of projects within the realm of geophysical exploration,
besides the ones we present here, \todo{Others coming to your mind?}e.g.
pyGIMLi \citep{CAG.17.Rucker} and Fatiando \citep{JOSS.18.Uieda}. Newer
\todo{wrong term, think about it}programming environments  have also brought
much simplifications in the installation process and have most importantly
simplified it for code developers to make their codes easily available and
installable. All our presented codes are in the Python ecosystem, and each code
can be installed with a single command (which includes downloading and
installation).

A second topic we want to address is verification. New codes are generally
verified against analytical solutions. However, analytical and semi-analytical
solutions only exist for very simple models. There is no possibility to check
how good a computed response is, expect for comparing the result from different
modellers. If different modellers using different discretisation and
implementations of Maxwell's equations yield the same result it can give
confidence in the accuracy. Providing these two models and results from four
different codes should give a benchmark with which new codes can be compared to
and validated with.

It is worth mentioning that there are many more 3D EM codes developed in other
fields than geophysics such as communication systems (antennas, radar,
satellites), medical imaging. There are various open-source codes in these
fields \todo{Add at least three references} too, e.g., Elmer
(\href{http://www.csc.fi/elmer}{csc.fi/elmer}, couldn't find a publication,
just a poster). While these codes could potentially be used for the same goals
as presented here we restrict our review to codes purpose-built, and ready
without further adjustments, for geophysical applications.

In the following section we introduce the codes under consideration. The
numerical result consider two cases. First a layered background with vertical
transverse isotropy containing three blocks, followed by a realistic, complex
3D model.

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}. All four
codes have their user-facing routines written in Python; all of them make heavy
use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy}
\citep{NM.20.Virtanen}. The four of them are “modern” open-source projects,
meaning that they not only come with an open-source license, but they are also
in an online-hosted version-control system with feedback and tracking
possibilities (raising issues, filing pull requests), have extensive online
documentation including many examples, have continuous integration to some
degree, and also importantly are all easily installable, without any user-side
compilations.

In the following a quick rundown of the codes. It is, however, beyond the scope
of this article to go into every detail of the different modellers, and we
refer to their documentations for more details. An overview comparison of the
different codes is given in Table~\ref{tbl:codecomparison}. All codes have in
common that they solve the\mycom[DW]{are we all using the weak form?} weak
formulation of Maxwell's equation in its differential form (DE) under the
quasistatic or\mycom[DW]{are we all using the diff. approx?} diffusive
approximation, hence neglecting displacement currents. The machines on which
the different codes were run are listed in Table~\ref{tbl:machines}, together
with the responsible operator.

\tabl[btp]{codecomparison}{All codes solve the weak form of Maxwell's equations
  in its differential form (DE) under the quasistatic approximation. Note that
  \emg3d is a solver on its own, while the other codes implement third-package
  solvers such as \texttt{PETSc}, \texttt{MUMPS}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
  \imycom[DW]{\texttt{PETSc} and \texttt{MUMPS} probably need citations too.}
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of computer and operating system (hardware and
software) on which the different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & PowerEdge R940 server; 144 Xeon Gold 6154 CPU @2.666 GHz; ~3 TB
            DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & Laptop; i7-6600U CPU@2.6 GHz x4; 16 GB of memory, Ubuntu 18.04
          & Dieter Werthmüller \\
  \petgem & Marenostrum4. Intel Xeon Platinum from Skylake generation; 2
            sockets Intel Xeon Platinum 8160 CPU with 24 cores each @2.10GHz
            for a total of 48 cores per node; 386 Gb DDR4 RAM per node; SuSE
            Linux Enterprise
          & Octavio Castillo-Reyes \\
  \simpeg & ???
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%

\imycom[DW]{
Each code should be briefly summarized by three points:\\
- Brief intro\\
- Main selling point\\
- Planned features\\
I think in a first round \textbf{everyone should just write what they want to
write and think is important for their code}. In a second round we should then
homogenize it, such that all codes have more or less the same length/weight and
also similar topic/layout.
}

@@

TODO (\texttt{FEniCS}; \cite{CSE.15.Alnaes})

@@

The 3D CSEM modeller \emg3d is a multigrid \citep{CMMP.64.Fedorenko} solver for
3D electromagnetic diffusion following \cite{GP.06.Mulder}, with tri-axial
electrical anisotropy, isotropic electric permittivity, and isotropic magnetic
permeability. The matrix-free solver can be used as main solver or as
preconditioner for one of the Krylov subspace methods implemented in SciPy, and
the governing equations are discretized on a staggered grid by the
finite-integration technique \citep{AEU.77.Weiland}, which is a finite-volume
generalization of a Yee grid. The code is written completely in Python using
the NumPy and SciPy stacks, where the most time- and memory-consuming parts are
sped up through jitted (just-in-time compiled) functions using Numba
\citep{LLVM.15.Lam}. The code computes the electric field due to an electric
source in the frequency-domain. Implemented are functions to obtain the
magnetic field due to an electric source; to obtain the electric and magnetic
fields due to a magnetic source; and to obtain time-domain responses.

The multigrid method is characterized by almost linear scaling both in terms of
runtime (CPU) and memory (RAM) usage, and it is therefore a comparably
low-memory consumption solver. It is also minimal in terms of requirements,
only NumPy, SciPy, and Numba are required, with Python 3.7+. The current
development is focused on adding inversion capabilities. A further plan is also
to implement a hook in order to use \emg3d as a solver within the \simpeg
framework.

@@

TODO

@@

TODO

\clearpage  % TODO: Just for now while drafting, remove later!
\section{Numerical Verifications}

@@

\cite{GJI.13.Miensopust} published a review of two workshops dealing with the
verification of magnetotelluric forward and inversion codes, hence a similar
exercise to what we do here for CSEM codes. The block model is a derivation of
the \emph{Dublin Test Model 1}, a model from their first workshop. We took the
same layout of the blocks but adjusted the dimensions and resistivities to a
typical, marine CSEM problem, as shown in Figure~\ref{fig:block-model}. We
additionally added a layered background with vertical transverse isotropy
(VTI).
%
\plot*{block-model}{width=.6\textwidth}{
  Sketch of the block model. The layered model consists of an air layer, a
  water layer, a thin top-layer followed by a thick, anisotropic background
  layer, and at the bottom a resistive basement layer. All three blocks are
  embedded in the thick background layer, which has VTI with
  $\lambda=\sqrt{2}$.
}
%

@@

Figure~\ref{fig:results-layered-0} shows in the top-row the inline responses
for offsets $r>0.5\,$km. As all codes compute the same we only show the
semi-analytical result for the real and imaginary amplitudes, and then in the
bottom row the relative error of the 3D codes. The results show that the codes
are literally yielding the same result, the relative error is for big parts
below 1\,\%, and for almost all parts below a few percent. Some differences are
due to the chosen discretizations. E.g., \emg3d is less precise close to the
200\,m long source. This could be improved by choosing smaller cell sizes, at
the cost of computation time. At large offsets, on the other hand, \emg3d seems
to do a very good job.
%
\plot*{results-layered-0}{width=.8\textwidth}{
   Comparison of the layered model for the inline receivers ($y=0\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   shows the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

@@

%
\tabl[btp]{comp-layered}{Comparison of used CPU and RAM and of number of cells
  used to discretize the layered model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#Cells \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

Three resistive blocks are added in the 3.3\,km-thick, anisotropic background
layer, as shown in Figure~\ref{fig:block-model}. They have resistivities of
$\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines),
$\rho=100\,\ohmm$ (thin plate, South-East), $\rho=500\,\ohmm$ (cube,
North-West). As there are no analytical solutions for this model we show the
normalized difference between different codes, instead of the relative error,
where the normalized difference of two responses $p$ and $q$ is given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200 \frac{|p - q|}{|p| + |q|}\ .
  \label{eq:nrmsd}
\end{equation}
%
\imycom[DW]{NRMSD: Comparing all codes to all would make the
figures too cluttered, and I don't think it would add a lot insights. My idea
is therefore to compare \emg3d/\simpeg (the two FV ones), \custem/\petgem (the
two FE ones), and \emg3d/\custem (a cross-comparison). What do you think?}

@@

The difference becomes only significant at offsets larger than 8\,km. However,
this is due to a practical difference in the meshing: This simple block model
is advantageous for FV codes, as it consists of rectangular blocks and
horizontal layers, simple geometric objects. While having a very thin, shallow
layer is straight-forward for FV codes, it requires many cells for a FE code.
The thin layer was therefore limited in its horizontal extent
to\mycom[DW]{Raphael, to what x/y?} ???. This is the reason for the increased
NRMSD at large offsets, and it has nothing to do with the accuracy of the
computation itself.
